LaMachine: A meta-distribution for NLP software

### 2018-09-10 docs/lamachine_extabstract_clarin2018.tex
[paper] updated Software section, less footnotes

LaMachine also aims to harmonise the metadata of all installed software, by converting metadata from upstream
repositories\footnote{i.e. the repositories where tool providers deposit their software} to the CodeMeta
standard\footnote{\url{https://codemeta.github.io/}, described in JSON-LD} \cite{codemeta,codemetar} where possible, or
encouraging software developers to provide their own. This in turn enables other tools to do proper service discovery
and provenance logging. Leveraging this metadata, LaMachine comes with a webserver that offers a portal website with
access and overview of all locally installed tools, including web services and web applications.

@@

As LaMachine was conceived initially\footnote{LaMachine exists since May 2015 and has been used extensively ever since by
numerous users} as the primary means of distribution of the software stack developed at CLST,
Radboud University Nijmegen, it includes a lot of our software. A full list goes beyond the scope of this overview; we
will merely mention some CLARIN-NL/CLARIAH-funded tools: ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format
for Linguistic Annotation, with assorted tools), FLAT (a web-based linguistic Annotation tool), PICCL (an OCR and
post-OCR correction pipeline) and CLAM. We already include some relevant software by the University of Groningen and VU
Amsterdam as well. Moreover, LaMachine incorporates a large number of renowned tools by external international parties,
offering most notably a mature Python environment\footnote{scipy, numpy, scikit-learn, matplotlib, nltk, spacy, pytorch,
keras, gensim, etc...}, but also R and Java.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use the Jupyter Notebook environment\footnote{\url{https://jupyter.org/}}, which offers a more graphical scripting environment
for e.g. Python and R that has gained great popularity in the scientific community.

>>>>

LaMachine also aims to harmonise the metadata of all installed software, by converting metadata from upstream
repositories\footnote{i.e. the repositories where tool providers deposit their software} to the CodeMeta
standard\footnote{\url{https://codemeta.github.io/}, described in JSON-LD} \cite{codemeta,codemetar} where possible, or
encouraging software developers to provide their own. This in turn enables other tools to do proper service discovery
and provenance logging.

Leveraging this metadata, LaMachine comes with a webserver that offers a portal website with
access and overview of all installed tools, including web services and web applications. It also comes with a
Jupyter Lab\footnote{\url{https://jupyter.org/}} environment which provides a web-based Integrated Development
Environment (IDE) for scripting in Python and R, web-based terminal access, and so-called \emph{notebooks} which mix
text and code and have gained great popularity in data science community nowadays.

@@

As LaMachine was conceived initially as the primary means of distribution of the software stack developed at CLST, Radboud University
Nijmegen, it includes a lot of our software. LaMachine exists since May 2015 and has been used extensively ever since
by numerous users. A full list of included software goes beyond the scope of this overview; we will merely mention
some CLARIN-NL/CLARIAH-funded tools: ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format for Linguistic
Annotation, with assorted tools), FLAT (a web-based linguistic Annotation tool), PICCL (an OCR and post-OCR correction
pipeline) and CLAM. We already include some relevant software by the University of Groningen (such as Alpino, a
dependency parser) and VU Amsterdam as well. Moreover, LaMachine incorporates a large number of renowned tools by
external international parties, offering most notably a mature Python environment with renowned scientific modules such
as scipy, numpy, scikit-learn, matplotlib, nltk, spacy, pytorch, keras, gensim, and many others, but also R, Java and tools
such as Stanford CoreNLP.

@@

We used a virtual machine (VM) image of LaMachine and we ran this 64-bits Linux-based VM inside another
VM with Windows Server 2012, provided to us by the health inspectorate for this project, in which we did have administrative
rights but no internet access. In terms of hardware we ran on a machine with 8 cores and  32GB internal memory available.
%using VirtualBox and Vagrnt on the Windows VM
LaMachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use
the Jupyter Notebook environment.
